---
title: "BarChart"
author: "Sumon Barua"
output:
  html_document:
    keep_md: true
    fig_width: 5
    fig_height: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Examples were taken from the book: 

* Graphics Cookbook: Practical Recipes for Visualizing Data
* Cookbook for R


Bar graph typically used to display:

* Numeric values (on the y-axis)
* Categories (on the x-axis)

####Heights of bars 

* The count of cases for each group - stat_count - stat="count"
* The value of a column - stat_identity - stat="identity"

```{r}
#install.packages("gcookbook")
library(gcookbook)
library(tidyverse)
pg_mean

ggplot(pg_mean, aes(x=group, y=weight)) + 
  geom_bar(stat="identity")

```

When x is a continuous (or numeric) variable vs discrete (categorical variable with factor())
```{r}
BOD

#continuous
ggplot(BOD, aes(x=Time, y=demand)) + 
  geom_bar(stat="identity")

#discrete
ggplot(BOD, aes(x=factor(Time), y=demand)) + 
  geom_bar(stat="identity")

```

####Colouring - 

* fill - to fill the bars
* colour - bar outline color
```{r}
pg_mean

ggplot(pg_mean, aes(x=group, y=weight)) +
  geom_bar(stat="identity", fill="lightblue", colour="black")

```

####Grouping bars together

* Map a variable to fill
* Use geom_bar(position="dodge")

```{r}
cabbage_exp


ggplot(cabbage_exp, aes(x=Date, y=Weight, fill = Cultivar )) +
  geom_bar(stat = "identity", position = "dodge")

```

####Custom colour palette
To set the colors, cale_fill_brewer() or scale_fill_manual() can be used
```{r}
cabbage_exp

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
   geom_bar(stat = "identity", position="dodge", colour="black") +
   scale_fill_brewer(palette="Pastel1")
```


####Bar graph with count
```{r}
diamonds

ggplot(diamonds, aes(x=cut)) +
   geom_bar(stat = "count")
```

####Custom colour
Custom colours can be used with help of scale_fill_manual
```{r}
upc <- subset(uspopchange, rank(Change)>40)
upc

ggplot(upc, aes(x= Abb , y = Change, fill = Region)) +
 geom_bar(stat="identity") +
 scale_fill_manual(values=c("#ef98aa", "#c2e3ec")) +
 xlab("State")

#Sort data
ggplot(upc, aes(x= reorder(Abb,Change) , y = Change, fill = Region)) +
 geom_bar(stat="identity") +
 scale_fill_manual(values=c("#ef98aa", "#c2e3ec")) +
 xlab("State")
```


####Bar graph with count
```{r}
diamonds

ggplot(diamonds, aes(x=cut)) +
   geom_bar(stat = "count")
```

####Positive and Negative Bars - Opsite direction

* stat = "identity"
* position = "identity"

```{r}
climate %>%
  filter(Source == "Berkeley" & Year >= 1900) %>%
  mutate(Status = Anomaly10y >= 0) %>%
  ggplot(aes(x = Year, y = Anomaly10y, fill = Status )) +
  geom_bar(stat = "identity", position = "identity")
```

####Proportional stacked bar graph
```{r}

prop_wt_tbl <- cabbage_exp %>%
  group_by(Date,Cultivar) %>%
  summarise(sw = sum(Weight)) %>%
  mutate(prop = sw/sum(sw)*100)

prop_wt_tbl

ggplot(prop_wt_tbl, aes(x=Date, y=prop, fill=Cultivar)) +
  geom_bar(stat="identity", colour="black") +
  scale_fill_brewer(palette="Pastel1")

```


####Labels to bar
```{r}
diamonds_cut_sum <- diamonds %>%
                  count(cut)

diamonds_cut_sum

ggplot(diamonds_cut_sum, aes(x=cut, y=n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label=n), vjust=-0.2) 

ggplot(diamonds_cut_sum, aes(x=cut, y=n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label=n), vjust=1.5, colour="white") 

##outward vs inward
```


####Cleveland Dot plot.

* By default, character vector will be ordered alphabetically. 
* By default, Factor would use the order defined in the factor levels.
* Re-ordering name based on values. ie. reorder(name, avg). This turns name is ordered factor.
* reorder() function will only order factor levels by one other variable


```{r}
top_records <- tophitters2001[1:25,]
top_records[, c("name", "lg", "avg")]


ggplot(top_records, aes(x=avg, y=name)) +
  geom_point(size=3)

ggplot(top_records, aes(x=avg, y=reorder(name, avg))) +
  geom_point(size=3) + # Use a larger dot
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour="grey60", linetype="dashed"))

#swap axis
ggplot(top_records, aes(x=reorder(name, avg), y=avg)) +
  geom_point(size=3) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle=60, hjust=1),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(colour="grey60", linetype="dashed"))


top_records$name[order(top_records$lg)]
top_records$name[order(top_records$lg, top_records$avg)]

#color dot based on group
ggplot(top_records, aes(x=avg, y=reorder(name, avg))) +
  geom_segment(aes(yend=name), xend=0, colour="grey50") + #line tail
  geom_point(size=3, aes(colour=lg)) +
  scale_colour_brewer(palette="Set1", limits=c("NL","AL")) +
  theme_bw() +
  theme(panel.grid.major.y = element_blank(), # No horizontal grid lines
        legend.position=c(1, 0.55), # Put legend inside plot area
        legend.justification=c(1, 0.5))

#facet_grid + and within group name sorted by avg
ggplot(top_records, aes(x=avg, y=reorder(name, avg))) +
  geom_segment(aes(yend=name), xend=0, colour="grey50") +
  geom_point(size=3, aes(colour=lg)) +
  scale_colour_brewer(palette="Set1", limits=c("NL","AL"), guide=FALSE) +
    theme_bw() +
    theme(panel.grid.major.y = element_blank()) +
    facet_grid(lg ~ ., scales="free_y", space="free_y")

```
