---
title: "BarChart"
author: "Sumon Barua"
output:
  html_document:
    keep_md: true
    fig_width: 5
    fig_height: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Examples were taken from the book: 

* Graphics Cookbook: Practical Recipes for Visualizing Data
* Cookbook for R


Bar graph typically used to display:

* Numeric values (on the y-axis)
* Categories (on the x-axis)

####Heights of bars 

* The count of cases for each group - stat_count - stat="count"
* The value of a column - stat_identity - stat="identity"

```{r}
#install.packages("gcookbook")
library(gcookbook)
library(tidyverse)
pg_mean

ggplot(pg_mean, aes(x=group, y=weight)) + 
  geom_bar(stat="identity")

```

When x is a continuous (or numeric) variable vs discrete (categorical variable with factor())
```{r}
BOD

#continuous
ggplot(BOD, aes(x=Time, y=demand)) + 
  geom_bar(stat="identity")

#discrete
ggplot(BOD, aes(x=factor(Time), y=demand)) + 
  geom_bar(stat="identity")

```

####Colouring - 

* fill - to fill the bars
* colour - bar outline color
```{r}
pg_mean

ggplot(pg_mean, aes(x=group, y=weight)) +
  geom_bar(stat="identity", fill="lightblue", colour="black")

```

####Grouping bars together

* Map a variable to fill
* Use geom_bar(position="dodge")

```{r}
cabbage_exp


ggplot(cabbage_exp, aes(x=Date, y=Weight, fill = Cultivar )) +
  geom_bar(stat = "identity", position = "dodge")

```

####Custom colour palette
To set the colors, cale_fill_brewer() or scale_fill_manual() can be used
```{r}
cabbage_exp

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
   geom_bar(stat = "identity", position="dodge", colour="black") +
   scale_fill_brewer(palette="Pastel1")
```


####Bar graph with count
```{r}
diamonds

ggplot(diamonds, aes(x=cut)) +
   geom_bar(stat = "count")
```

####Custom colour
Custom colours can be used with help of scale_fill_manual
```{r}
upc <- subset(uspopchange, rank(Change)>40)
upc

ggplot(upc, aes(x= Abb , y = Change, fill = Region)) +
 geom_bar(stat="identity") +
 scale_fill_manual(values=c("#ef98aa", "#c2e3ec")) +
 xlab("State")

#Sort data
ggplot(upc, aes(x= reorder(Abb,Change) , y = Change, fill = Region)) +
 geom_bar(stat="identity") +
 scale_fill_manual(values=c("#ef98aa", "#c2e3ec")) +
 xlab("State")
```


####Bar graph with count
```{r}
diamonds

ggplot(diamonds, aes(x=cut)) +
   geom_bar(stat = "count")
```

####Positive and Negative Bars - Opsite direction

* stat = "identity"
* position = "identity"

```{r}
climate %>%
  filter(Source == "Berkeley" & Year >= 1900) %>%
  mutate(Status = Anomaly10y >= 0) %>%
  ggplot(aes(x = Year, y = Anomaly10y, fill = Status )) +
  geom_bar(stat = "identity", position = "identity")
```
