---
title: "Explore"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import Libraries
```{r eval = FALSE}
library(tidyverse)
library(lubridate)
library(stringr)
library(readr)
library(ggplot2)
library(forcats)

```





## Global variables

This section will list all the global variables needed for this sesstion.

```{r}
ds_path <- ""
```


## Loading file
```{r eval = FALSE}
ds <- readr::read_csv(ds_path)
```

## Additional class
Assigning aditional class for better dataset management
```{r eval = FALSE}
ds<- tbl_df(ds)
```
##Basic info
```{r eval = FALSE}
dim(ds)
class(ds)
sapply(ds,class)
names(ds)
str(ds)
summary(ds)
```

##Sample
```{r eval = FALSE}
head(ds)
tail(ds)

#Random sample
dplyr::sample_n(ds, 5)
```


## Meta data cleansing
Sometimes it is convenient to map all variable names to lowercase, and remove additional character
```{r eval = FALSE}
names(ds) <- stringr::str_to_lower(names(ds)) 
names(ds) <- stringr::str_replace_all(stringr::str_to_lower(names(mainDs)) , "[ /.]", "_") 
```


## Identify empty columns
```{r eval = FALSE}
# 100% missing values
ds_info_values_missing <- sapply(ds, function(x) sum(is.na(x)))
ds_info_columns_empty <- names(which(ds_info_values_missing == nrow(ds)))
ds[,ds_info_columns_empty] <- NULL

# Example 70% missing values
s_info_columns_empty_percent <- names(which(ds_info_values_missing >= 0.70*nrow(ds)))
ds[,ds_info_columns_empty_percent] <- NULL
ds_info_columns_to_keep <- setdiff()
``` 
##Conversion
Converting string to numeric, and also remove $ and ,. Below **ks_prepare_amount** function is documented in the common function document.
```{r eval = FALSE}
#sale_price is example field
ds <- ks_prepare_amount(ds, "sale_price")
```

Converting string to date
```{r eval = FALSE}
#sale_date is example field
ds$sale_date <- lubridate::dmy(ds$sale_date)
```
