---
title: "Predictive Analysis"
author: "Sumon Barua"
output:
  html_document:
  keep_md: true
fig_width: 5
fig_height: 3
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Examples were taken from the book: 
  
  * Graphics Cookbook: Practical Recipes for Visualizing Data
  * Cookbook for R


```{r}
#install.packages("tidyverse")
library(tidyverse)

```


####Correlation Analysis

* Range of correlation coefficient -> -1 to +1
* Perfect positive relationship    -> +1
* Perfect Negative relationship    -> -1
* No linear relationship           -> 0
* Good relationship                -> >+-0.85
* Correlation is positive when the values increase together
* Correlation is negative when one decreases when other value increases
* Scatter plot helps to identify correlation visually
* Correlation works well when the relationship is a straight line
* Sometimes calculation may not pick up relationship but visuals can
* Correlation is not causation - Correlation does not mean that one thing causes other.

```{r}
diamonds

ggplot(diamonds, aes(x=carat, y=price)) +
  geom_point()

```

####Scatter Plot

* Good relationship



####Simple linear regression

* An extension of correlation.
* Best fitting straight line for a scatter plot between two variables

