---
title: "Scallter Plot"
author: "Sumon Barua"
output:
  html_document:
  keep_md: true
fig_width: 5
fig_height: 3
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Examples were taken from the book: 
  
  * Graphics Cookbook: Practical Recipes for Visualizing Data
  * Cookbook for R


```{r}
#install.packages("tidyverse")
library(tidyverse)
library(gcookbook)
```

####Scatter Plot

* Relationship between two continuous variables
* Each observation is represented by a point

```{r}
head(heightweight[,c("ageYear", "heightIn")])

ggplot(heightweight, aes(x=ageYear, y=heightIn)) +
  geom_point()

```

####Different shape+size for point

* Default is solid circles -> 16
* Default size is 2
* i.e. shape=21
* i.e. size=1.5
* Shape 19 is also a solid circle looks smoother

```{r}

ggplot(heightweight, aes(x=ageYear, y=heightIn)) +
  geom_point(shape=19, size=1.5)

```

####Grouping data point by a variable - (factor) (color/shape)

* colour=variable_name -> part of aes
* shape=variable_name -> part of aes
* Group variable must be categorical/factor/character vector
* Numeric group variable : Numeric vector needs to converted to factor 
* http://ggplot2.tidyverse.org/reference/scale_brewer.html
* http://colorbrewer2.org/



```{r}
head(heightweight[, c("sex", "ageYear", "heightIn")])

ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
  geom_point()

ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex)) +
  geom_point()

ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex, colour=sex)) +
  geom_point() +
  scale_shape_manual(values=c(1,2)) +
  scale_colour_brewer(palette="Set1")
```


####Grouping data point by a continuous variable - (numerical) (color/shape) 

```{r}

head(heightweight[, c("sex", "ageYear", "heightIn", "weightLb")])

ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
  geom_point()

```

